---
title: "Vis_II"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(haven)
```


```{r weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature scatterplot", 
    x = "Minimum Temp (C)", 
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15ยบ C", "0", "15"), 
limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

Looking at colors:

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(h = c(100, 300)) 
```

### I would look at color pallets that are already made. Maybe use the virdis color pallet. 

```{r}
ggp_temp_plot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
```


## Themes
```{r}
ggp_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  viridis::scale_color_viridis(discrete = TRUE)
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom")
# moving legend position
```

```{r}
ggp_scatterplot +
  theme_bw() +
  # background is black and white
  theme(legend.position = "bottom")
```

Order matters...

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_bw()
```

using theme bw puts everything back; resets; so be aware of order. 

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_minimal()
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_classic()
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  ggthemes::theme_excel()
```

Learning assesment
New Scatterplot

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se=FALSE) +
  viridis::scale_color_viridis(discrete = TRUE) +
  labs(
    title = "Seasonal Variation in Max Temp",
    x = "Date",
    y = "Maxiumum temperature (C)",
    size = "Percipitation"
  ) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```


Extra bonus stuff in `ggplot`

USe different data sets in different `geom`s

```{r}
central_park_df =
  weather_df |>
  filter(name == "CentralPark_NY")

molokai_df =
weather_df |>
  filter(name == "Molokai_HI")

molokai_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```


## Multiple panels

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill=name)) +
  geom_density() +
  facet_grid(. ~ name)
```


run patch work library before this: 

```{r}
ggp_tmax_tmin = 
  weather_df |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3)

ggp_tmax_density =
  weather_df |>
  ggplot(aes(x=tmax, fill= name)) +
  geom_density(alpla=.3)

ggp_tmax_date =
  weather_df |>
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point() +
  geom_smooth(se=FALSE)

(ggp_tmax_tmin + ggp_tmax_density)/ ggp_tmax_date
```


## Data manipulation

```{r}
weather_df |>
  mutate(name=fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |>
  ggplot(aes(x=name, y=tmax, fill=name)) +
  geom_violin(alpha=.5)
```

Pulse data next:

```{r}
pulse_df=
  read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  # have to create variables before making the plot.
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to= "visit",
    values_to= "bdi_score",
    names_prefix = "bdi_score_"
  ) |>
  mutate(visit = ifelse(visit == "bl", "00m", visit))
# whats happening in parenthesis: take bl replace w 00m and otherwise leave alone. the box plots will now be in order of time (before numbers came before letters)

pulse_df |>
  ggplot(aes(x=visit, y=bdi_score)) +
  geom_boxplot()
```





